Progress Report for CSS 586: \\ Modeling Latent Patterns in Music

### 2021-04-25 papers/progress/alex/progress-alex.tex
add note-seq dependency for reading MIDI files into Python
add some midi deps and paper work

\begin{abstract}
This report explores recent research in modeling music with deep learning and
provides a progress report on a project to train a generative model of classical
music on the MusicNet dataset.
\end{abstract}

@@

Depending on the task, machine learning models of music may be trained on the
audio signal itself, either in a time domain or a frequency domain
representation, or they may be trained on a digital symbolic representation of
music, the most common of which is MIDI (Musical Instrument Digital Interface)
notation. MIDI is an encoding of music as streams of bytes in one or more tracks
or channels, each representing a sequence of 128 possible pitch values, along
with timing, pressure and instrument values. A music transcription model may
convert an audio signal into MIDI, which can easily be converted into other
symbolic representations such as sheet music, while a synthesizer model can
convert MIDI representations into audio signals.

@@

A paper from Boston University describes an effort to combine the symbolic and
waveform approaches to music modeling, by training an LSTM to learn melodic
structure of different styles of music, then providing generations from this
model as conditioning inputs to a WaveNet-based raw audio generator
\cite{manzelli_conditioning_2018}.

\section{Planned Methods}

\section{Progress}

>>>>

\begin{abstract}
This report explores recent research in modeling music with deep learning and
provides a progress report on a course project to train a generative model of
classical music on the MusicNet dataset.
\end{abstract}

@@

Depending on the task, machine learning models of music may be trained on the
audio signal itself, either in a time domain or a frequency domain
representation, or they may be trained on a digital symbolic representation of
music, the most common of which is MIDI (Musical Instrument Digital Interface)
notation. MIDI is an encoding of music as streams of bytes in one or more tracks
or channels, each representing a sequence of 128 possible pitch values, along
with timing, pressure and instrument values. A music transcription model may
convert an audio signal into MIDI, which can easily be converted into other
symbolic representations such as sheet music for human performers to read from,
while a synthesizer model can convert MIDI representations into audio signals.

@@

Prior work points out that the division between symbolic music notes and the
sounds of music is analogous to the division between symbolic language and
utterances in speech, which may inspire ideas for combining the two approaches
\cite{hawthorne2019enabling}. A paper from Boston University describes an effort
to combine the symbolic and waveform approaches to music modeling, by training
an LSTM to learn melodic structure of different styles of music, then providing
generations from this model as conditioning inputs to a WaveNet-based raw audio
generator \cite{manzelli_conditioning_2018}.

\section{Planned Methods and Progress}

\subsection{Data Preprocessing}

This project will focus on the generative modeling of symbolic music using MIDI
data, because of the advantages of symbolic music models in representing
long-term structure in musical compositions to produce generations with coherent
use of repetition over long time scales.

\subsection{Model Fitting}

We will explore several modeling approaches to generating symbolic music:

\begin{itemize}
  \item{Sliding window sequence prediction with RNNs (LSTM/GRU)}
  \item{Sliding window sequence prediction with Transformers}
  \item{Latent space interpolation with Sequential VAEs}
  \item{Latent space interpolation with Sequential GANs}
\end{itemize}

\subsection{Model Evaluation}

Evaluation of generative models is challenging because there is no equivalent of
an accuracy metric like what is used in supervised learning. Generative models
are typically evaluated using a combination of qualitative metrics whereby human
judges rate the quality of the generated examples (essentially a Turing test),
and quantitative metrics that assess the differences in the parametric
distributions of generated and real examples. Yang and Lerch (2020) proposes a
set of metrics informed by music theory, for probabilistically evaluating how
similar the generations are to known sample distributions of real music
\cite{yang_evaluation_2020}. These metrics include counts, ranges, histograms
and transition matrices of pitches and note lengths, then the Kullback-Leibler
divergence and overlapping area of the probability density functions are use to
compare against known reference distributions per musical genre
\cite{yang_evaluation_2020}. Due to the cost and time requirements associated
with designing a human subjects experiment, we plan to utilize this quantitative
approach to generation quality assessment.
